### 2019-07-25 0_abstract.tex
Changes to trim word count.

\begin{abstract}
% 148 words
We present Task Bench, a \emph{parameterized} benchmark designed to
explore the performance envelopes of parallel and distributed
programming systems under a variety of application scenarios. Task
Bench lowers the barrier to benchmarking multiple programming systems by making
the implementation for a given system orthogonal to the benchmarks
themselves: all benchmarks constructed with Task Bench automatically
run on all Task Bench implementations, and vice versa. Furthermore,
Task Bench's parameterization allows a wide variety of benchmark
scenarios that distill the key characteristics of larger applications.

>>>>

\begin{abstract}
% 148 words
We present Task Bench, a \emph{parameterized} benchmark designed to
explore the performance of parallel and distributed
programming systems under a variety of application scenarios. Task
Bench lowers the barrier to benchmarking multiple programming systems by making
the implementation for a given system orthogonal to the benchmarks
themselves: every benchmark constructed with Task Bench
runs on every Task Bench implementation. Furthermore,
Task Bench's parameterization allows a wide variety of benchmark
scenarios that distill the key characteristics of larger applications.

### 2019-07-25 1_introduction.tex
Changes to trim word count.

We present Task Bench, a parameterized benchmark designed for exploring the performance
behaviors of parallel and distributed programming systems under a
variety of conditions.  The key property of Task Bench is that it completely separates
the system-specific implementation from the core implementation
of the benchmarks themselves.
In all benchmarks we know of, the effort to implement $n$ benchmarks on $m$
systems is $\mathcal{O}(mn)$.  Task Bench's design reduces this work to $\mathcal{O}(m + n)$,
enabling dramatically more systems and benchmarks to be explored for the same amount of programming
effort.  New benchmarks created with Task Bench
immediately run on all systems, and new systems immediately run all
benchmarks.

@@

Adding a new programming system to Task Bench involves  implementing a set of standard
services, such as executing a task or a data transfer.  This interface is simple enough to 
be implemented with reasonable effort to a level of quality sufficient for comparative
studies. Task Bench provides a core library that encapsulates the functionality
shared by all system implementations, which not only reduces the implementation effort
required but also makes it much easier to achieve truly apples-to-apples comparisons
between systems.

@@

These results point to a natural question:   How does one predict whether the performance of a runtime
system will in fact be good enough for a particular application?  In trying to answer this question,
we realized that \emph{weak} and \emph{strong} scaling, by far the most commonly
reported performance metrics, do not directly characterize the performance of the underlying
programming system. Weak scaling can hide arbitrary amounts of runtime
system overhead by using sufficiently large problem sizes, and strong
scaling does not separate runtime system overhead from costs
(such as increased communication) that scale with the number of nodes when
using progressively larger portions of a machine.

To characterize the contribution of a programming model's runtime
system to application performance, and as an example of the novel studies that can be done
with Task Bench, we introduce a new metric called
\emph{minimum effective task granularity} (METG). Intuitively, for a given
workload, METG(50\%) is the smallest task granularity that maintains
at least 50\% efficiency, meaning that the application achieves at
least 50\% of the higest performance (in FLOP/s, B/s, or another
application-specific measure) that can be achieved on a given
machine. The efficiency bound in METG is a key innovation over
previous approaches, such as \emph{tasks per second} (TPS), that fail
to quantify the amount of useful work peformed (if tasks are
non-empty~\cite{Canary16, Armstrong14}) or to perform any useful work at all (if tasks are empty~\cite{LegionTracing18}).

METG captures the essence of what is important in a
weak or strong scaling study, the behavior at the limit of
scalability. For weak scaling, METG(50\%) corresponds directly to the
smallest problem size that can be weak-scaled with 50\%
efficiency. For strong scaling, METG(50\%) can be used to compute the
scale at which performance can be expected to dip below 50\%
efficiency.  We note that METG(50\%) for a given runtime system will
vary with the particular application and the underlying hardware---i.e., METG(50\%)
is not a constant for a given system, but we will see that individual systems have
a characteristic range of METG(50\%) values and that there is additional insight
in the reasons that METG can vary.

@@

The paper is organized as follows: Section~\ref{sec:task-bench}
describes the design of Task Bench, Section~\ref{sec:implementation}
details the implementation of Task Bench in 15 programming systems,
Section~\ref{sec:metg} defines the METG(50\%) metric and its
relationship to quantities of interest to application developers,
Section~\ref{sec:evaluation} provides a comprehensive evaluation of
Task Bench on up to 256 Haswell nodes of the Cori
supercomputer~\cite{Cori}, Section~\ref{sec:related-work} relates this
work to previous efforts, and Section~\ref{sec:conclusion} concludes.

>>>>

We present Task Bench, a parameterized benchmark designed for exploring the performance
of parallel and distributed programming systems under a
variety of conditions.  The key property of Task Bench is that it completely separates
the system-specific implementation from the core implementation
of the benchmarks themselves.
In all benchmarks we know of, the effort to implement $n$ benchmarks on $m$
systems is $\mathcal{O}(mn)$.  Task Bench's design reduces this work to $\mathcal{O}(m + n)$,
enabling dramatically more systems and benchmarks to be explored for the same amount of programming
effort.  New benchmarks created with Task Bench
immediately run on all systems, and new systems that implement the Task Bench interface immediately run all
benchmarks.

@@

Adding a new programming system to Task Bench involves  implementing a set of standard
services, such as executing a task or a data transfer.  This interface is simple enough to 
be implemented with reasonable effort to a level of quality sufficient for comparative
studies. Task Bench provides a core library that encapsulates the functionality
shared among systems, which not only reduces the implementation effort
required but also makes it much easier to achieve truly apples-to-apples comparisons
between systems.

@@

This result points to a natural question:   How does one predict whether the performance of a runtime
system will in fact be good enough for a particular application?  In trying to answer this question,
we realized that \emph{weak} and \emph{strong} scaling, by far the most commonly
reported performance metrics, do not directly characterize the performance of the underlying
programming system. Weak scaling can hide arbitrary amounts of runtime
system overhead by using sufficiently large problem sizes, and strong
scaling does not separate runtime system overhead from costs
(such as increased communication) that scale with the number of nodes when
using progressively larger portions of a machine.

To characterize the contribution of a programming model's runtime
system to application performance, and as an example of the novel studies that can be done
with Task Bench, we introduce a new metric called
\emph{minimum effective task granularity} (METG). Intuitively, for a given
workload, METG(50\%) is the smallest task granularity that maintains
at least 50\% efficiency, meaning that the application achieves at
least 50\% of the higest performance (in FLOP/s, B/s, or another
application-specific measure) that can be achieved on a given
machine. The efficiency bound in METG is a key innovation over
previous approaches, such as \emph{tasks per second} (TPS), that fail
to consider the amount of useful work peformed (if tasks are
non-empty~\cite{Canary16, Armstrong14}) or to perform any useful work at all (if tasks are empty~\cite{LegionTracing18}).

METG captures the essence of what is important in a
weak or strong scaling study, the behavior at the limit of
scalability. For weak scaling, METG(50\%) corresponds to the
smallest problem size that can be weak-scaled with 50\%
efficiency. For strong scaling, METG(50\%) can be used to compute the
scale at which performance can be expected to dip below 50\%
efficiency.  We note that METG(50\%) for a given runtime system will
vary with the particular application and the underlying hardware---i.e., METG(50\%)
is not a constant for a given system, but we will see that individual systems have
a characteristic range of METG(50\%) values and that there is additional insight
in the reasons that METG can vary.

@@

The paper is organized as follows: Section~\ref{sec:task-bench}
describes the design of Task Bench, Section~\ref{sec:implementation}
discusses the implementation of Task Bench in 15 programming systems,
Section~\ref{sec:metg} defines the METG(50\%) metric and its
relationship to quantities of interest to application developers,
Section~\ref{sec:evaluation} provides a comprehensive evaluation of
Task Bench on up to 256 Haswell nodes of the Cori
supercomputer~\cite{Cori}, Section~\ref{sec:related-work} relates this
work to previous efforts, and Section~\ref{sec:conclusion} concludes.

### 2019-07-25 3_implementation.tex
More text and fiddle with figure placement.
Changes to trim word count.

In all cases, members of the respective programming systems' teams
were consulted during the development and evaluation of the
corresponding Task Bench implementations. Where assistance was provided, the insights were invaluable
in helping to ensure that we provide the highest level of quality in the Task
Bench implementations for each system.

>>>>

In all cases, members of the respective programming systems' teams
were consulted during the development and evaluation of the
corresponding Task Bench implementations. Where assistance was provided, the insights were invaluable
in helping to ensure that we provide the highest quality implementations for each system.

### 2019-07-25 4_metg.tex
Changes to trim word count.

METG(50\%) is the intersection of this curve with 50\% efficiency, as
shown by the red, dashed lines in Figure~\ref{fig:flops-mpi} and
\ref{fig:efficiency-mpi}. At 50\% efficiency, MPI achieves an average
task granularity of
4.6 \textmu{}s, thus the METG(50\%) of MPI is 4.6 \textmu{}s for this configuration of
Task Bench. We use 50\% because that is generally an acceptable level
of efficiency in practice, and values above 90\% can misrepresent the
performance of some systems (see
Section~\ref{subsec:peak-performance-and-efficiency}).

>>>>

METG(50\%) is the intersection of this curve with 50\% efficiency, as
shown by the red, dashed lines in Figure~\ref{fig:flops-mpi} and
\ref{fig:efficiency-mpi}. At 50\% efficiency, MPI achieves an average
task granularity of
4.6 \textmu{}s, thus the METG(50\%) of MPI is 4.6 \textmu{}s in this
configuration. We use 50\% because that is generally an acceptable level
of efficiency in practice, and values above 90\% can misrepresent the
performance of some systems (see
Section~\ref{subsec:peak-performance-and-efficiency}).

### 2019-07-25 5_evaluation.tex
More text and fiddle with figure placement.
Changes to trim word count.

\input{e1_flops}
\input{e2_efficiency}
\input{e7_bytes}
\input{e4_radix}

@@

The vast majority of systems achieve or nearly achieve peak FLOP/s. Some
systems reserve a number of cores (usually 1 or 2) for internal
runtime usage; these systems take a minor hit in peak FLOP/s compared
to systems which share all cores between the application and runtime. Some of the
higher overhead systems struggle to achieve peak FLOP/s, though in most cases
the curves suggest that performance would continue to improve if we
were to run larger problem sizes. Unfortunately, the excessive
computational cost of running such tests makes this prohibitively
expensive. For example, the Spark job in
this configuration ran for over 6 hours.

Figure~\ref{fig:efficiency} shows efficiency (as a percentage of
peak FLOP/s) plotted against task granularity. This plot is
used to calculate METG(50\%), as described in Section~\ref{sec:metg}. The
red, dashed line shows the point where 50\% efficiency is achieved.
In most cases, task granularity asymptotes prior to the 50\% efficiency point,
though some systems continue to improve as efficiency continues
to drop. Accounting for this effect is one of the main arguments in
favor of METG with a reasonable efficiency threshold instead of
measuring task scheduling throughput of empty tasks
(effectively METG(0\%)). Measuring performance using empty tasks can
reward implementation strategies, such as devoting nearly 100\% of
system resources to the runtime system, that make no sense for real
applications.

@@

Not all cores are required to saturate memory bandwidth. As a result,
systems that reserve one or more cores are able to hit 100\% of peak
performance, unlike with compute-bound kernels. Otherwise the
performance curves are similar. We are currently investigating the
performance of Charm++, Regent and X10.

@@

\input{e5_imbalance}

@@

\input{e6_communication}

>>>>

\input{e1_flops}
\input{e2_efficiency}

@@

The vast majority of systems achieve or nearly achieve peak FLOP/s. Some
systems reserve a number of cores (usually 1 or 2) for internal
runtime usage; these systems take a minor hit in peak FLOP/s compared
to systems which share all cores between the application and runtime. Some of the
higher overhead systems struggle to achieve peak FLOP/s, though in most cases
the curves suggest that performance would continue to improve if we
were to run larger problem sizes. Unfortunately, the excessive
computational cost of running such tests makes this prohibitively
expensive. For example, the Spark job in
this case ran for over 6 hours.

\input{e7_bytes}
\input{e4_radix}

Figure~\ref{fig:efficiency} plots efficiency (as a percentage of
peak FLOP/s) against task granularity. As described in Section~\ref{sec:metg}, this plot is
used to calculate METG(50\%). The
red, dashed line shows the point where 50\% efficiency is achieved.
In most cases, task granularity asymptotes prior to the 50\% efficiency point,
though some systems continue to improve as efficiency continues
to drop. Accounting for this effect is one of the main arguments in
favor of METG with a reasonable efficiency threshold instead of
measuring task scheduling throughput of empty tasks
(effectively METG(0\%)). Measuring performance using empty tasks can
reward implementation strategies, such as devoting nearly 100\% of
system resources to the runtime system, that make no sense for real
applications.

@@

Not all cores are required to saturate memory bandwidth. As a result,
systems that reserve one or more cores are able to hit 100\% of peak
performance, unlike with compute-bound kernels. We are currently investigating the
performance of Charm++, Regent and X10.

@@

@@

\input{e5_imbalance}
\input{e6_communication}
